---
title: "The hypogen package"
author: "Kosmas Hench"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
library(rtracklayer)
devtools::load_all("~/Desktop/hyplo_Rlab/hypogen")

```


## About the package:

The **hypogen** package helps with population genetic analysis based on the hamlet reference genome (*Hypoplectrus puella*). It contains one template for genome wide plots (`geom_hypo_LG()`) and one template to set statistics into the context of the genome annotation (`hypo_annotation_baseplot()`).

In addition to this the package contains a few helpers for a consistent plot layout.
More *artistic* aspects of hamlet analysis plots are provided by the [**hypoimg**](www.github.com/k-hench/hypoimg) package.

## The basic genome wide plot

Ploting a certain statistic over the whole genome is a frequent issue. An example for this is the running average of *F~ST~*.

Assume we have a dataset containing to variables measure along the genome:

```{r create_fake_genome_wide_data}
test_df <- tibble(x=rep(seq(1,max(hypo_karyotype$GEND),length.out = 500),2)) %>%
  mutate(CHROM = hypo_which_CHROM(x),grp = rep(c('sin','cos'), each = 500),
         y = ifelse(grp=='sin',sin(x/200000000*2*pi),cos(x/200000000*2*pi+.2))+rnorm(1000))
test_df %>% head
```

```{r plot_genome_wide,fig.width=10,fig.height=6, out.width="100%"}
ggplot() +
  geom_hypo_LG() +
  geom_line(data=test_df,aes(x=x,y=y,col=CHROM)) +
  scale_fill_hypo_LG_bg() +
  scale_x_hypo_LG() +
  scale_color_hypo_LG() +
  facet_grid(grp~.)+
  theme_hypo()
```

## The annotation plot

```{r create_fake_annotation_data}
XR <- c(0,100000)
test_df_2 <- tibble(x=rep(seq(XR[1],XR[2],length.out = 200),2)) %>%
  mutate(window = rep(c('bold(hyp_sin)','bold(hyp_cos)'), each = 200),
         y = ifelse(window == 'bold(hyp_sin)',
                    sin(x/XR[2]*20*pi),
                    cos(x/XR[2]*20*pi+.2)) + rnorm(400))
```

```{r plot_annotation,fig.width=10,fig.height=8, out.width="100%"}
hypo_annotation_baseplot(searchLG = "LG03", xrange = XR,
                         genes_of_interest = 'SUN1_4') +
  geom_line(data = test_df_2, aes(x = x, y = y, color = window)) +
  facet_grid(window~.,scales='free_y',
             switch = 'y',labeller = label_parsed,
             as.table = TRUE) +
  scale_color_manual(values = c(hypo_clr_LGs[c(1,24,10,15,20)]),
                     guide = FALSE) +
  scale_fill_manual(values = hypo_clr_LGs[c(10,15,20)], guide = FALSE) +
  scale_x_continuous(name = "Hypo AnnoPlot Title",
                     expand = c(0,0),position = 'top') +
  theme_hypo() +
  theme_hypo_anno_extra()
```
