---
title: "The hypogen package"
author: "Kosmas Hench"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hypogen intro}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{css, echo=FALSE}
img {
  border: 0;
}
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

devtools::load_all("~/Desktop/hypo_Rlab/hypogen/")

```

## About the package:

The **hypogen** package helps with population genetic analysis based on the hamlet reference genome (*Hypoplectrus puella*). It contains two functions to import population genetic statistics as produced by e.g. [VCFtools](https://vcftools.github.io/index.html) (`hypo_import_snps()` & `hypo_import_windows()`) in the context of the hamlet genome. Furthermore it provides one template for genome wide plots (`geom_hypo_LG()`) and one template to set statistics into the context of the genome annotation (`hypo_annotation_baseplot()`).

In addition to this the package contains a few helpers for a consistent plot layout.
More *artistic* aspects of hamlet analysis plots are provided by the [**hypoimg**](www.github.com/k-hench/hypoimg) package.

## Installation

Using the **devtools** package, **hypogen** can be installed directly from github:

```{r install, eval = FALSE}
devtools::install_github("k-hench/hypogen")
```

## Importing Data
```{r import_data, warning = FALSE, message = FALSE}
# SNP data
file_snps <- system.file("extdata", "example.weir.fst.gz", package = "hypogen")
data_snps <- hypo_import_snps(file = file_snps, gz = TRUE) %>% 
  mutate(window = 'bolditalic(F[ST])')

# Window data
file_windows <- system.file("extdata", "example.windowed.weir.fst.gz", package = "hypogen")
data_windows <- hypo_import_windows(file = file_windows, gz = TRUE) %>% 
  mutate(window = 'bolditalic(F[ST])')
```

## The basic genome wide plot

Ploting a certain statistic over the whole genome is a frequent issue. An example for this is the running average of *F~ST~*.

Assume we have a dataset containing to variables measure along the genome:

```{r create_fake_genome_wide_data}
test_df <- tibble(x=rep(seq(1,max(hypo_karyotype$GEND),length.out = 500),2)) %>%
  mutate(CHROM = hypo_which_CHROM(x),grp = rep(c('sin','cos'), each = 500),
         y = ifelse(grp=='sin',sin(x/200000000*2*pi),cos(x/200000000*2*pi+.2))+rnorm(1000))
test_df %>% head
```

```{r plot_genome_wide,fig.width=10,fig.height=6, out.width="100%"}
ggplot() +
  geom_hypo_LG() +
  geom_line(data=test_df,aes(x=x,y=y,col=CHROM)) +
  scale_fill_hypo_LG_bg() +
  scale_x_hypo_LG() +
  scale_color_hypo_LG() +
  facet_grid(grp~.)+
  theme_hypo()
```

## The annotation plot

```{r create_fake_annotation_data}
XR <- c(10000,50000)
test_df_2 <- tibble(x=rep(seq(XR[1],XR[2],length.out = 200),2)) %>%
  mutate(window = rep(c('bold(hyp_sin)','bold(hyp_cos)'), each = 200),
         y = ifelse(window == 'bold(hyp_sin)',
                    sin(x/XR[2]*20*pi),
                    cos(x/XR[2]*20*pi+.2)) + rnorm(400))
```

```{r plot_annotation,fig.width=10,fig.height=8, out.width="100%"}
hypo_annotation_baseplot(searchLG = "LG01", xrange = XR,
                         genes_of_interest = "TSEN34",
                         genes_of_sec_interest = "Serp1") +
  geom_point(data = data_snps, aes(x = POS, y = WEIR_AND_COCKERHAM_FST),
            color = 'lightgray') +
  geom_line(data = data_windows, aes(x = POS, y = WEIGHTED_FST),
            color = viridis::inferno(5)[4]) +
  coord_cartesian(xlim = XR) +
  facet_grid(window~.,scales='free_y',
             switch = 'y',labeller = label_parsed) +
  scale_color_manual(values = c('black','gray',viridis::inferno(3)),
                     guide = FALSE) +
  scale_fill_manual(values = viridis::inferno(3),
                    guide = FALSE) +
  scale_x_continuous(name = "Hypo AnnoPlot Title",
                     expand = c(0,0),position = 'top') +
  theme_hypo() +
  theme_hypo_anno_extra()
```

------------------------

<center>
<img src = "logo.svg" width = 20% />
</center>
