---
title: "genotype frequencies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{genotype frequencies}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{css, echo=FALSE}
img {
  border: 0;
}
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library("hypogen")
library("svglite")
```


## Disclaimer:

This used to work - unfortunately {ggtern} seems to be broken at tihs moment.
This page will be updated if {ggtern} gets fixed...

## Preparation:

To be able to plot genotytpe frequencies, we need a genotype table where genotypes
are encoded as 0,1, & 2 (homozygous reference allele, heterozygous and homozygous alternative allele).

This can be easily generated from a **vcf file** using **VCFtools**:

```sh
vcftools \
  --gzvcf input.vcf.gz \
  --out  genotype_table \
  --012
```

Unfortunately the **vcftools** output has the wrong orientation (samples as rows, sites as columns).
So before we reed the genotypes into **R** they are transformed using **awk**:

```sh
awk '
{
    for (i=1; i<=NF; i++)  {
        a[NR,i] = $i
    }
}
NF>p { p = NF }
END {
    for(j=1; j<=p; j++) {
        str=a[1,j]
        for(i=2; i<=NR; i++){
            str=str" "a[i,j];
        }
        print str
    }
}' genotype_table.012 | gzip > genotype_table.012.trans.txt.gz

```
## R setup

To efficiently plot the genotypes frequencies we make use of the **ggtern** package as well as of the **tidyverse** (and of the **hypogen** package):

```{r load_libs, eval = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggtern)
```

## Loading the data into R

The package **hypogen** provides the function `hypo_import_genotype_freq()` which can read
the transformed genotypee table and directly computes genotype frequencies:

```{r data_import, message = FALSE, warning = FALSE}
file_genotypes <- system.file("extdata", "genotype_table.012.trans.txt.gz", package = "hypogen")

genotype_freqs <- hypo_import_genotype_freq(file_genotypes)
```

This table can be directly plottet using **gggtern**.
As a reference we also make use of the **hypogen** function `hypo_hwe()` to indicate Hardy-Weinberg
genotype frequencies:

```{r plot, eval = FALSE,fig.width=10,fig.height=6, out.width="100%", message = FALSE, warning = FALSE}
ggplot() +
  coord_tern()+
  geom_point(data = genotype_freqs,
             aes(x = AA, y = Aa, z = aa ),
             col = hypo_clr_LGs[18]) +
  geom_path(data = hypo_hwe(n = 100),
            aes(x = AA, y = Aa, z = aa ),
            color ='black',alpha = .6)
```

------------------------

```{r, echo = FALSE, fig.asp = 1, out.height = "150pt", out.width = "150pt", dev = "svglite", fig.align = "center"}
library(ggforce)
hypoimg:::hypo_logo()
```
